Reproducibility of COVID-19 pre-prints

### 2021-07-22 outputs/paper/paper-copy.Rmd
Minor updates to paper, update slides to match paper

---
title: "Reproducibility of COVID-19 pre-prints"
author: 
  - Annie Collins^[University of Toronto.]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca.]
thanks: "We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Collins thanks CANSSI Ontario and Monica Alexander for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, medRxiv, and SocArXiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, 79 per cent of those on medRxiv, and 85 per cent of those on SocArXiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

published_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("bioRxiv" = 2, "medRxiv" = 2, "arXiv" = 2, "SocArXiv" = 2))
```

@@

To examine the influence of the COVID-19 pandemic on open science practices in during the pandemic, we analyzed pre-prints posted between January and December 2019 from each of the four repositories using the same process as was done for the COVID-19-related pre-prints. Since medRxiv was founded in June 2019, all pre-prints posted in the latter half of 2019 were analyzed (a total of 913). For all other repositories, a random sample of 1,200 was taken from all non-COVID-19-related pre-prints posted in the relevant date range.

@@

# TODO: position = "dodge" not working
all_2019 %>% 
  mutate(month = floor_date(date, "month")) %>% 
  # filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month, fill = type)) + 
  geom_histogram(stat="count", position = "dodge2") +
  labs(title = "Pre-prints posted per month, 2019",
       subtitle = "Total of 150,018 pre-prints on arXiv and 31,752 on bioRxiv, 913 pre-prints on \n medRxiv and 1,614 on SocArXiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~type,
             scales = "free_y")
```

@@

The other concern is the adverse selection issue caused by meeting the open science aims of sharing code and data. Authors that share their data and code open their work up to criticism. If authors who make their data and code available make similar mistakes to authors who choose to not publish their data and code, it is more likely that the mistake would not be noticed in the case where data and code were not published. The current system is biased against those who follow best practice. Indeed, @mcguinness2021descriptive advocate '(s)trict editorial policies that mandate data sharing'.

>>>>

---
title: "Reproducibility of COVID-19 pre-prints"
author: 
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca.]
  - Annie Collins^[University of Toronto.]
thanks: "We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Collins thanks CANSSI Ontario and Monica Alexander for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, medRxiv, and SocArXiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, 79 per cent of those on medRxiv, and 85 per cent of those on SocArXiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

published_open_status %>%
  count(repo, published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  mutate(Published = ifelse(published == 0, "No", "Yes"), `Proportion with neither`= round((`Neither`)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  select(Published, Both, Neither, `Open code`, `Open data`, `Proportion with neither`) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("arXiv" = 2, "bioRxiv" = 2, "medRxiv" = 2, "SocArXiv" = 2))
```

@@

To examine the influence of the COVID-19 pandemic on open science practices during the pandemic, we analyzed pre-prints posted between January and December 2019 from each of the four repositories using the same process as was done for the COVID-19-related pre-prints. Since medRxiv was founded in June 2019, all pre-prints posted in the latter half of 2019 were analyzed (a total of 913). For all other repositories, a random sample of 1,200 was taken from all non-COVID-19-related pre-prints posted in the relevant date range.

@@

all_2019 %>% 
  mutate(month = floor_date(date, "month")) %>% 
  # filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month, fill = type)) + 
  geom_histogram(stat="count", position = "dodge2") +
  labs(title = "Pre-prints posted per month, 2019",
       subtitle = "Total of 150,018 pre-prints on arXiv and 31,752 on bioRxiv, 913 pre-prints on \n medRxiv and 1,614 on SocArXiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~type,
             scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

@@

The other concern is the adverse selection issue caused by meeting the open science aims of sharing code and data. Authors that share their data and code open their work up to criticism. If authors who make their data and code available make similar mistakes to authors who choose to not publish their data and code, it is more likely that the mistake would not be noticed in the case where data and code were not published. The current system is biased against those who follow best practice. Indeed, @mcguinness2021descriptive advocate for "(s)trict editorial policies that mandate data sharing."
